Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-10-31 manuscript.tex
added degen and goodman footnote

In this paper we asked whether linking hypothesis and number of
response options available to participants in truth value judgment tasks
affects inferred implicature rates. The results presented here suggest
they do. A linking assupmtion that considered the highest point on the
scale literal and any lower point pragmatic (Weak link) resulted in
higher implicature rates in tasks with 4 or 5 response options compared
to the standard two options. A linking hypothesis that considered the
lowest point on the scale pragmatic and any higher point literal (Strong
link) reported lower implicature rates in tasks with 4 or 5 options
compared to the standard two options. The results suggest that the
choice of linking hypothesis is a crucial analytical step that can
significantly impact the conclusions drawn from truth value judgment
tasks. In particular, there is danger for pragmatic ability to be both
under- and overestimated.

While the binary truth value judgement task avoids the analytic decision
between Strong and Weak linking hypothesis, the results reported here
suggest that binary tasks can also underestimate participants' pragmatic
competence. In binary tasks, participants are often given the lowest and
highest points on a scale (\enquote{wrong} vs. \enquote{right}) and are
asked to report pragmatic infelicities using the lowest point (e.g.
\enquote{wrong}). The study reported here showed that on trials with
true but pragmatically infelicitous descriptions, participants often
avoided the lowest point on the scale if they were given more
intermediate options. Even though the option \enquote{wrong} was
available to participants in all tasks, participants in tasks with
intermediate options chose it less often. In computing implicature rate,
this pattern manifested itself as a decrease in implicature rate under
the Strong link when more response options were provided, and an
increase in implicature rate under the Weak link when more response
options were provided. These observations are in line with Katsos and
Bishop (2011)\enquote{s argument that pragmatic violations are not as
severe as semantic violations and participants do not penalize them as
much. Providing participants with only the extreme ends of the scale
(e.g.~wrong/right, false/true) when pragmatic violations are considered
to be of an intermediate nature risks misrepresentation of participants}
pragmatic competence. It further suggests that in studies that use
binary tasks to investigate response-contingent processing, proportions
of \enquote{literal} responses may be a composite of both literal and
pragmatic underlying interpretations that just happen to get mapped
differently onto different response options by participants.

@@

We contrast the above view of implicature and its behavioral reflexes
with an alternative linking hypothesis. Recent developments in the field
of probabilistic pragmatics have demonstrated that pragmatic production
and comprehension can be captured within the Rational Speech Act (RSA)
framework (Bergen, Levy, \& Goodman, 2016; Degen, Franke, \& Jäger,
2013; Degen, Tessler, \& Goodman, 2015; Frank \& Goodman, 2012; Franke
\& Jäger, 2016; Goodman \& Frank, 2016; Goodman \& Stuhlmüller, 2013;
Kao, Wu, Bergen, \& Goodman, 2014; Qing \& Franke, 2015; Scontras,
Degen, \& Goodman, 2017). Much in the spirit of Gricean approaches to
pragmatic competence, the RSA framework takes as its point of departure
the idea that individuals are rational, goal-oriented communicative
agents, who in turn assume that their interlocutors similarly behave
according to general principles of cooperativity in communication. Just
as in more traditional Gricean pragmatics, pragmatic inference and
pragmatically-cooperative language production in the RSA framework are,
at their core, the product of counterfactual reasoning about alternative
utterances that one might produce (but does not, in the interest of
cooperativity). However, the RSA framework explicitly and quantitatively
models cooperative interlocutors as agents whose language production and
comprehension is a function of Bayesian probabilistic inference
regarding other interlocutors' expected behavior in a discourse context.

@@

We take inspiration from the RSA approach and treat participants'
behavior in our experimental tasks as the result of a soft-optimal
pragmatic speaker in the RSA framework. That is, following Degen and
Goodman (2014), we proceed on the assumption that behavior on sentence
verification tasks, such as truth value judgment tasks, is best modeled
as a function of an individual's mental representation of a cooperative
interlocutor (\(S_1\) in the language of RSA) rather than of a pragmatic
listener who interprets utterances (\(P_{L_1}\)). In their paper, Degen
\& Goodman argue that sentence verification tasks are relatively more
sensitive to contextual manipulations (such as manipulation of the
Question Under Discussion) than are sentence interpretation tasks, and
that this follows if sentence interpretation tasks -- but not sentence
verification tasks -- require an additional layer of counterfactual
reasoning about the intentions of a cooperative speaker.

>>>>

In this paper we asked whether linking hypothesis and number of
response options available to participants in truth value judgment tasks
affects inferred implicature rates. The results presented here suggest
they do. A linking assumption that considered the highest point on the
scale literal and any lower point pragmatic (Weak link) resulted in
higher implicature rates in tasks with 4 or 5 response options compared
to the standard two options. A linking hypothesis that considered the
lowest point on the scale pragmatic and any higher point literal (Strong
link) reported lower implicature rates in tasks with 4 or 5 options
compared to the standard two options. The results suggest that the
choice of linking hypothesis is a crucial analytical step that can
significantly impact the conclusions drawn from truth value judgment
tasks. In particular, there is danger for pragmatic ability to be both
under- and overestimated.

While the binary truth value judgement task avoids the analytic decision
between Strong and Weak linking hypothesis, the results reported here
suggest that binary tasks can also underestimate participants' pragmatic
competence. In binary tasks, participants are often given the lowest and
highest points on a scale (\enquote{wrong} vs. \enquote{right}) and are
asked to report pragmatic infelicities using the lowest point (e.g.
\enquote{wrong}). The study reported here showed that on trials with
true but pragmatically infelicitous descriptions, participants often
avoided the lowest point on the scale if they were given more
intermediate options. Even though the option \enquote{wrong} was
available to participants in all tasks, participants in tasks with
intermediate options chose it less often. In computing implicature rate,
this pattern manifested itself as a decrease in implicature rate under
the Strong link when more response options were provided, and an
increase in implicature rate under the Weak link when more response
options were provided. These observations are in line with Katsos and
Bishop (2011)'s argument that pragmatic violations are not as
severe as semantic violations and participants do not penalize them as
much. Providing participants with only the extreme ends of the scale
(e.g.~wrong/right, false/true) when pragmatic violations are considered
to be of an intermediate nature risks misrepresentation of participants'
pragmatic competence. It further suggests that in studies that use
binary tasks to investigate response-contingent processing, proportions
of \enquote{literal} responses may be a composite of both literal and
pragmatic underlying interpretations that just happen to get mapped
differently onto different response options by participants.

@@

We contrast the above view of implicature and its behavioral reflexes
with an alternative linking hypothesis. Recent developments in the field
of probabilistic pragmatics have demonstrated that pragmatic production
and comprehension can be captured within the Rational Speech Act (RSA)
framework (Bergen, Levy, \& Goodman, 2016; Degen, Franke, \& Jäger,
2013; Degen, Tessler, \& Goodman, 2015; Frank \& Goodman, 2012; Franke
\& Jäger, 2016; Goodman \& Frank, 2016; Goodman \& Stuhlmüller, 2013;
Kao, Wu, Bergen, \& Goodman, 2014; Qing \& Franke, 2015). Much in the spirit of Gricean approaches to
pragmatic competence, the RSA framework takes as its point of departure
the idea that individuals are rational, goal-oriented communicative
agents, who in turn assume that their interlocutors similarly behave
according to general principles of cooperativity in communication. Just
as in more traditional Gricean pragmatics, pragmatic inference and
pragmatically-cooperative language production in the RSA framework are,
at their core, the product of counterfactual reasoning about alternative
utterances that one might produce (but does not, in the interest of
cooperativity). However, the RSA framework explicitly and quantitatively
models cooperative interlocutors as agents whose language production and
comprehension is a function of Bayesian probabilistic inference
regarding other interlocutors' expected behavior in a discourse context.

@@

We take inspiration from the RSA approach and treat participants'
behavior in our experimental tasks as the result of a soft-optimal
pragmatic speaker in the RSA framework. That is, following Degen and
Goodman (2014), we proceed on the assumption that behavior on sentence
verification tasks such as truth value judgment tasks, is best modeled
as a function of an individual's mental representation of a cooperative
\change{speaker} (\(S_1\) in the language of RSA) rather than of a pragmatic
listener who interprets utterances (\(P_{L_1}\)).\footnote{\change{Degen and Goodman (2014) argue that sentence verification is more plausibly construed as a production task rather than as an interpretation task because participants, unlike in natural language comprehension, are provided with the ground truth about the state of the world that a speaker is describing. Thus, participants are in essence being asked to assess the quality of a speaker's utterance. In contrast, Degen and Goodman argue, true interpretation tasks are characterized by the listener inferring what the state of the world is that the speaker is describing, for instance by selecting from one of multiple interpretation options.}} In their paper, Degen
\& Goodman \change{show} that sentence verification tasks are relatively more
sensitive to contextual \change{features like} the
Question Under Discussion than are sentence interpretation tasks, and
that this follows if sentence interpretation tasks -- but not sentence
verification tasks -- require an additional layer of counterfactual
reasoning about the intentions of a cooperative speaker.
